---
layout: page
title: Individual Growth
subtitle: Other VBGF Parameterizations
author: Derek H. Ogle
css: "/css/codebox.css"
csl: ../american-fisheries-society.csl
bibliography: ../IFARSupplement.bib
output: 
  html_document: 
    fig_width: 5
    self_contained: no
---
```{r echo=FALSE, eval=FALSE}
##############################################################
# == BEGIN -- NOT SHOWN IN SUPPLEMENT, FOR PRINTING ONLY =====
# First line renders an appropriate HTML file for the webpage
# Second line makes the script file
# RUN BOTH MANUALLY (following using Knit HTML button)
setwd("C:/aaaWork/Web/GitHub/IFAR/supplements/growth")
source("../../rhelpers/IFARhelpers.R")
modHTML("OtherVBGFParams",need2render=FALSE)
if (require(FSA)) purl2("OtherVBGFParams.Rmd",topnotes="User must set working directory appropriately.")
# == END -- NOT SHOWN IN SUPPLEMENT, FOR PRINTING ONLY =====
##############################################################
```

The von Bertalanffy Growth Function (VBGF) was introduced and methods for fitting the model were illustrated in Chapter 12 of the [*Introductory Fisheries Analyses with R*](http://derekogle.com/IFAR/) (IFAR) book.  The VBGF used there is from @BevertonHolt1957 and is called the "Typical" parameterization of the VBGF.  The VBGF can be expressed in alternative forms that are called parameterizations.  The most common parameterizations of the VBGF may be seen with `vbModels()` from `FSA`.

Different parameterizations do not fit the data differently.  In other words, all parameterizations of the VBGF have the same number of parameters and will produce the same predicted values and fitted model plots.  However, different parameterizations will have different model parameters that are either of interest to the fisheries scientist or behave better with the model fitting algorithms.  For example, the parameterizations of @vonBertalanffy1938 and @Mooijetal1999 both use $L_{0}$, which is the mean length at time 0 (i.e., at "hatch").  Thus, either of these parameterizations would be used if the $L_{0}$ parameter was of biological interest.  In contrast, the parameterizations of @Schnute1981 and @Francis1988 are either largely or wholly based on expected values which are generally less correlated.  Thus, if one wants to avoid the problems associated highly correlated parametes (as discussed in the **IFAR** book) and is not interested in the parameters of the other parameterizations, then either of these two parameterizations may be used.

The parametarization of @Francis1988 will be fit in this supplement as an illustration of fitting other than the "Typical" parameterization that was demonstrated in Chapter 12 of the **IFAR** book.  Other supplements demonstrate how to  [fit other growth functions](OtherGrowthFuns.html) and [use alternative nonlinear model fitting algorithms](OtherNLAlgorithms.html).

\ 

### Required Packages for this Supplement
Functions used in this supplement require the packages shown below.

```{r echo=FALSE, results='hide', message=FALSE, warnings=FALSE}
##############################################################
# == BEGIN -- NOT SHOWN IN SUPPLEMENT, FOR PRINTING ONLY =====
# 
# Setup of knitr
source("../../rhelpers/knitr_setup.R")
# declare packages used
rqrd <- c("FSA","dplyr","nlstools","captioner","knitr")
# setup figure, table, and equation captioning
library(captioner)
figcaps <- captioner(prefix="Figure")
figcaps("","")
# == END -- NOT SHOWN IN SUPPLEMENT, FOR PRINTING ONLY =====
##############################################################
```
```{r cache=FALSE, results='hide', message=FALSE}
library(FSA)
library(dplyr)
library(nlstools)
```

### Data Used in this Supplement
  
The male Black Drum data ([view](https://github.com/droglenc/FSAdata/blob/master/data-raw/BlackDrum2001.csv), [download](https://raw.githubusercontent.com/droglenc/FSAdata/master/data-raw/BlackDrum2001.csv), [meta-data](http://derekogle.com/fishR/data/data-html/BlackDrum2001.html)) used in the **IFAR** book are also used in this supplement.
```{r}
bdmf <- read.csv("BlackDrum2001.csv") %>%
  select(-c(spname,day,weight)) %>%
  filterD(sex %in% c("male","female"),otoage<50)
bdm <- filterD(bdmf,sex=="male")
headtail(bdm)
```

\ 

## Fitting Other VBGF Parameterizations

An R function for predicting length from a given age may be created by the user.  However, similar to what was demonstrated in the main chapter for the "Typical" VBGF,  `vbFuns()` from `FSA` may be used to create such a function for each parameterization displayed in `vbModels()`.  A brief explanation of the parameters is given if `msg=TRUE` is included in `vbFuns()`.  A function for the "Francis" parameterization is defined with `vbFuns()` below.  Note that values for the three parameters may be included as a vector in `L1=` and that the youngest and oldest ages may be included as a vector in `t1=`.

```{r}
( vbF <- vbFuns("Francis") )
```

The three parameters in the "Francis" parameterizations are the expected (i.e., mean) lengths at three ages.  Two of the ages are generally the youngest and oldest (or very close to those) ages in the data.  The third age is, by definition, halfway between these two ages.  Thus, the parameters will be the expected lengths at a young, old, and intermediate age.  In this example, I have chosen to estimate expected lengths for the youngest and oldest ages in the data.
```{r}
( ages <- range(bdm$otoage) )
```

Starting values are obtained with `vbStarts()` as described in the main chapter.  However, with the "Francis" parameterization, the ages that define the youngest and oldest age must also be provided in `ages2use=`.

```{r}
( svF <- vbStarts(tl~otoage,data=bdm,type="Francis",ages2use=ages) )
```

The model is fit and parameter estimates with likelihood profile are obtained as described in the book, with the exception that the vector that defines the youngest and oldest ages must be supplied when `vbF()` is used.  Bootstrapped confidence intervals and predictions may also be obtained as described in the book.
```{r}
nlsF <- nls(tl~vbF(otoage,L1,L2,L3,t1=ages),data=bdm,start=svF)
cbind(Ests=coef(nlsF),confint(nlsF))
bootF <- nlsBoot(nlsF)
cbind(Ests=coef(nlsF),confint(bootF))
predict(bootF,vbF,t=3,t1=ages)
```

Note that the intra-parameter correlations are relatively low for this parameterization.  Additionally, the scale of the parameters in the "Francis" parameterization are usually similar.  Both of these characteristics may aid model convergence.
```{r}
summary(nlsF,correlation=TRUE)
```

\ 

--------------------------------------------------------------

```{r echo=FALSE, results='asis'}
reproInfo(rqrdPkgs=rqrd,out="markdown",links=c(Script="OtherVBGFParams.R",RMarkdown="OtherVBGFParams.Rmd"))
```

--------------------------------------------------------------

\ 

## References
