---
layout: page
title: Age Comparisons
subtitle: Pymatuning Sanctuary Walleye
author: Derek H. Ogle
output: 
  html_document: 
    fig_width: 5
    self_contained: no
---
```{r echo=FALSE}
source("../rhelpers/IFARhelpers.R")
```

[Kocovsky and Carline (2000)](http://www.tandfonline.com/doi/abs/10.1577/1548-8675%282000%29020%3C1044%3AACOMFE%3E2.0.CO%3B2) examined methods to estimate the age of the unexploited population of [Walleye](https://en.wikipedia.org/wiki/Walleye) in [Pymatuning Sanctuary](http://www.fish.state.pa.us/stocking/warmcool/pyma_spring03.htm) (PA).  [Scales](https://en.wikipedia.org/wiki/Fish_scale), dorsal spines, and [otoliths](https://en.wikipedia.org/wiki/Otolith) were extracted from fish sampled in trap nets during the spawning run in March and April of 1997. Two readers independently examined each structure without knowledge of fish length or sex.  Only consensus ages for each structure are recorded in `r addDataLinks("WalleyePS")`.

1. Use tabular, graphical, and statistical methods to describe any apparent bias between age estimates from scales and otoliths.
1. Describe any apparent bias between age estimates from scales and spines.
1. Describe any apparent bias between age estimates from spines and otolith.
1. Does the sex of the fish impact any of the conclusions made in the three previous questions.

---
```{r echo=FALSE, results="asis"}
exercise_footer("AgeComp_PSWalleye")
```
