---
layout: page
title: Individual Growth
subtitle: Alaskan Slimy Sculpins B
author: Derek H. Ogle
output: 
  html_document: 
    fig_width: 5
    self_contained: no
---
```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/IFAR/exercises"); source("../rhelpers/IFARhelpers.R")
fnm <-"AKSlimySculpins_Growth_B"; modHTML(fnm)
```

The total length (mm) and otolith age of [Slimy Sculpin](https://en.wikipedia.org/wiki/Slimy_sculpin) (*Cottus cognatus*) captured in the [Arctic Long-Term Ecological Research](http://ecosystems.mbl.edu/ARC/) area were recorded in [this CSV file](SculpinALTER.csv) (these data are also available in `SculpinALTER` from `FSAdata`.

Use these data, and results from [this exercise](AKSlimySculpins_Growth_A.html), to answer the following questions.

1. Plot TL versus age.  Observe the "shape" of the data (do the results look linear or like a von Bertalanffy growth curve, is there an obvious asymptote, are young fish well represented, how variable are lengths within ages).
1. Fit the von Bertalanffy's original parameterization (see `vbModels()`).
    1. Write the VBGF with parameters replaced by their estimated values.
    1. Carefully interpret the meaning of each parameter?
    1. Predict the mean TL, with 95% confidence interval, for an age-3 Slimy Sculpin.  How does this prediction compare to the prediction you made using the typical VBGF in [this exercise](AKSlimySculpins_Growth_A.html)?
    1. Plot TL versus age and superimpose the best-fit VBGF.  Comment on model fit.
    1. Construct a residual plot.  Comment on model fit.
    1. How does the estimate of $L_{\infty}$ and $K$ from fitting this parameterization compare to that from the typical VBGF fit in [this exercise](AKSlimySculpins_Growth_A.html).  Explain your observation.

---
```{r echo=FALSE, results="asis"}
exercise_footer(fnm)
```
