---
layout: page
title: Age-Length Keys
subtitle: Lake Erie Freshwater Drum C
author: Derek H. Ogle
output: 
  html_document: 
    fig_width: 5
    self_contained: no
---
```{r echo=FALSE}
source("../rhelpers/IFARhelpers.R")
```

[Bur (1984)](http://www.sciencedirect.com/science/article/pii/S0380133084718065) examined the population dynamics of [Freshwater Drum](https://en.wikipedia.org/wiki/Freshwater_drum) (*Aplodinotus grunniens*) in  [Lake Erie](https://en.wikipedia.org/wiki/Lake_Erie) in the late 1970s.  In one part of his study, he measured the total length (TL) of all 1577 drum sampled and extracted [scales](https://en.wikipedia.org/wiki/Fish_scale) for age estimation from a proportionate sample from each 10 mm length interval.  The length and age data are recorded in `r addDataLinks("FWDrumLE2")`.

1. Separate the observed data into age- and length-samples.  Construct an **observed** age-length key.  [This step would have been accomplished in [this exercise](LEFWrum_ALK_A.html).]
1. Use the "classical" method to estimate the age distribution (with standard errors) for all sampled fish.
    1. How many fish are estimated to be age 3?
    1. How many fish are estimated to be age 8?
    1. Plot the age distribution for all fish.
1. Use the "classical" method to estimate the mean length-at-age (with standard deviations) for all sampled fish.
    1. What is the mean TL of age-4 fish?
    1. Plot the length-at-age with the mean length-at-age superimposed for all fish.
1. Compare your results to someone else's results (or repeat the steps above).  Did you both get the *exact* same results?  Why or why not?  If not, how different were they?
1. Compare your results using the "classical" method here to your results from using the Isermann and Knight (2005) method in [this exercise](LEFWDrum_ALK_B.html) (assuming that you completed that exercise).

[Continue with these data here](LEFWDrum_ALK_C.html).

---
```{r echo=FALSE, results="asis"}
exercise_footer("LEFWDrum_ALK_B")
```
