---
layout: page
title: Abundance from Mark-Recapture Data
subtitle: Wisconsin YOY Walleye
author: Derek H. Ogle
output: 
  html_document: 
    fig_width: 5
    self_contained: no
---
```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/IFAR/exercises"); source("../rhelpers/IFARhelpers.R")
fnm <-"MarkRecap_WIYOYWalleye"; modHTML(fnm)
```

[Mraz (1968)](http://digicoll.library.wisc.edu/cgi-bin/EcoNatRes/EcoNatRes-idx?type=header;pview=hide;id=EcoNatRes.DNRBull40) examined the population dynamics of young-of-the-year (YOY) [walleye](https://en.wikipedia.org/wiki/Walleye) (*Sander vitreus*) in Pine Lake, an inland lake in Wisconsin.  In fall 1962, YOY walleye were captured, marked, and returned to the lake on five sampling dates.  On each date, the number of fish caught, the number of caught fish that were previously marked, and the number of marked fish returned to the lake were recorded.  The results from these sampling efforts are shown in the table below.

| Sample |  &nbsp;&nbsp;&nbsp;Caught&nbsp;&nbsp;&nbsp;  | &nbsp;&nbsp;&nbsp;Recaptured&nbsp;&nbsp;&nbsp; | &nbsp;&nbsp;&nbsp;Returned&nbsp;&nbsp;&nbsp; |
|:------:|:--------:|:----------:|:--------:|
|    1   |    321   |     --     |    321   |
|    2   |    412   |     45     |    412   |
|    3   |    178   |     55     |    178   |
|    4   |    415   |     93     |    415   |
|    5   |    367   |    113     |     --   |

Use these results to answer the questions below.

1. Estimate the initial population size, with 95% confidence interval, of YOY Walleye in Pine Lake.  Carefully interpret the result.
1. Comment on the validity of the assumptions for the model you used to estimate the size of the population.

---
```{r echo=FALSE, results="asis"}
exercise_footer(fnm)
```
