---
layout: page
title: Recruitment
subtitle: Escanaba Lake (WI) Walleye (B)
author: Derek H. Ogle
output: 
  html_document: 
    fig_width: 5
    self_contained: no
---

```{r echo=FALSE}
source("../rhelpers/IFARhelpers.R")
```

[Hansen *et al.*](http://www.tandfonline.com/doi/abs/10.1577/1548-8675%281998%29018%3C0764%3AFAROWI%3E2.0.CO%3B2?journalCode=ujfm20) (1998) modeled recruitment variation of age-0 [Walleye](https://en.wikipedia.org/wiki/Walleye) (*Sander vitreus*) in [Escanaba Lake](http://dnr.wi.gov/lakes/lakepages/LakeDetail.aspx?wbic=2339900&page=facts) (Wisconsin) to determine factors regulating their abundance.  Specifically, they examined the abundance of age-5 and older Walleye (spawning population), variation in May water temperatures, and abundance of 152.4 mm total length and longer [Yellow Perch](https://en.wikipedia.org/wiki/Yellow_perch) (*Perca flavescens*) on the abundance of age-0 Walleye.  These data are available in [this CSV file](data/WalleyeEL.csv) (these data are also available in `WalleyeEL` from `FSAdata`).

Load these data into R to answer the following questions. [Note that this is an extension of this [recruitment exercise](Recruitment_ELWalleye_A.html).]

1. Fit the Ricker stock-recruitment function with a multiplicative error structure to age-5 and older and age-0 abundance data. [Note that this would have been completed in this [recruitment exercise](Recruitment_ELWalleye_A.html).]
1. Determine if adding the variation in May water temperatures or the abundance of Yellow Perch significantly improves the predictive power of the model.
1. If either of these two variables improved the predictive power of the model then answer the following.
    1. Express the best model by replacing parameters with their estimated values.
    1. Specifically describe the effect of the other variables on the recruitment of age-0 Walleye.
    1. Do you feel that this model provides a clear explanation for the variability in recruitment of age-0 Walleye in Escanaba Lake?  Explain.

---
```{r echo=FALSE, results="asis"}
exercise_footer("Recruitment_ELWalleye_B")
```
