---
layout: page
title: Weight-Length
subtitle: Trout Lake Yellow Perch A
author: Derek H. Ogle
output: 
  html_document: 
    fig_width: 5
    self_contained: no
---
```{r echo=FALSE}
source("../rhelpers/IFARhelpers.R")
```

Researchers for the [North Temperate Lakes Long-Term Ecological Research](https://lter.limnology.wisc.edu/) program have collected fish from a number of lakes in Wisconsin since 1981.  The total lengths and weights of [Yellow Perch](https://en.wikipedia.org/wiki/Yellow_perch) (*Perca flavescens*) captured from [Trout Lake](https://lter.limnology.wisc.edu/researchsite/trout-lake) from 1981-2006 are recorded in `r addDataLinks("YPerchTL")`.  Use these data to answer the following questions.  [*Note that these data came from [this website](https://lter.limnology.wisc.edu/dataset/north-temperate-lakes-lter-fish-lengths-and-weights-1981-current).  See the site for more information about the sampling protocol that produced these data.*]

Load these data, isolate Yellow Perch captured with only fyke nets in 2005, and answer the following questions.

1. Complete [this Plotting exercise](YPTroutLake_Plotting.html).
    1. Describe the relationship between weight and length.
    1. Describe the relationship between log-transformed weight and length.
1. Compute the weight-length relationship with an appropriate linear regression.
    1. Plot the results (data and the fitted relationship) on both the transformed and raw scales.  Comment on the fit.
    1. Construct a residual plot.  Comment.
    1. Express your results as an equation on the transformed scale.
    1. Express your results as an equation on the raw scale.
    1. Carefully interpret the meaning of the slope of the weight-length relationship.
    1. Is there statistical evidence for isometric or allometric growth?

Continue with these data for [weight-length comparisons](YPTroutLake_WL_B.html).

---
```{r echo=FALSE, results="asis"}
exercise_footer("YPTroutLake_WL_A")
```
