---
layout: page
title: Weight-Length
subtitle: Trout Lake Yellow Perch B
author: Derek H. Ogle
output: 
  html_document: 
    fig_width: 5
    self_contained: no
---
```{r echo=FALSE}
source("../rhelpers/IFARhelpers.R")
```

Researchers for the [North Temperate Lakes Long-Term Ecological Research](https://lter.limnology.wisc.edu/) program have collected fish from a number of lakes in Wisconsin since 1981.  The total lengths and weights of [Yellow Perch](https://en.wikipedia.org/wiki/Yellow_perch) (*Perca flavescens*) captured from [Trout Lake](https://lter.limnology.wisc.edu/researchsite/trout-lake) from 1981-2006 are recorded in `r addDataLinks("YPerchTL")`.  Use these data to answer the following questions.  [*Note that these data came from [this website](https://lter.limnology.wisc.edu/dataset/north-temperate-lakes-lter-fish-lengths-and-weights-1981-current).  See the site for more information about the sampling protocol that produced these data.*]

Load these data, isolate Yellow Perch captured with only beach seines and fyke nets in 1997, and answer the following questions.  [Note that this is a continuation of [this exercise](YPTroutLake_WL_A.html).]

1. Complete [this Plotting exercise](YPTroutLake_Plotting.html).
    1. Qualitatively assess if the weight-length relationship differs between the two years.
1. Fit a model that will allow you to determine if there is a statistically signficant difference in the weight-length relationship between the two years.
    1. Construct a residual plot for this model.  Comment.
    1. Is there a statistically signficant difference in the weight-length relationship between the two years?  Provide evidence for your findings and be very specific with your conclusions.
    1. Without fitting separate regressions for the two sample years, express the weight-length relationships on the raw scale for both years (i.e., write two specific equations).
    1. Construct a plot that illustrates your findings.

---
```{r echo=FALSE, results="asis"}
exercise_footer("YPTroutLake_WL_B")
```
