---
title: "Backcalculation of Previous Lengths"
author: "Derek H. Ogle"
output: html_document
---

:::{.callout-note}
The `RFishBC` package has functions to acquire growth increment data from images of ageing structures and to ease back-calculation computations. Please see [the vignettes for `RFishBC`](https://fishr-core-team.github.io/RFishBC/) for a more thorough description of back-calculating previous lengths.
:::

&nbsp;

@francis_back-calculation_1990 defined *back-calculation* as "*... a technique that uses a set of measurements made on a fish at one time to infer its length at an earlier time or times. Specifically, the dimensions of one or more marks in some hard part of the fish, together with its current body length, are used to estimate its length at the time of formation of each of the marks. ... The marks are generally annual rings associated with growth checks, ... .*" Thus, backcalculation is the reconstruction of the length of a fish at previous ages from measurements made on calcified structures.

# Setup
## Packages
Functions used in this supplement require the following packages.

```{r}
#| results: hide
library(FSA)      # for peek(), headtail(), sumtable()
library(dplyr)    # for filter(), select(), mutate(), arrange(), group_by(), summarize()
```

Also note that single functions from the `tidyr` and `stringr` packages are used with `::` so that the whole packages are not loaded here.

## Data
All analyses in this supplement use the West Bearskin Lake (MN) Smallmouth Bass (*Micropterus dolomieu*) data from @weisberg_using_1993 found in [SMBassWB](https://fishr-core-team.github.io/FSA/reference/SMBassWB.html).^[Download data with CSV link on the linked page.] The analysis in this supplement will focus on those fish captured in 1990. Three variables that were constant and not used in this analysis and three measurement variables that corresponded to ages that did not exist after reducing to only fish captured in 1990 were removed to save space.

```{r}
#| eval: false
SMBassWB <- read.csv("https://raw.githubusercontent.com/fishR-Core-Team/FSA/main/data-raw/SMBassWB.csv")
```
```{r}
#| echo: false
## work-around until the FSA package website is updated
data(SMBassWB,package="FSA")
```

```{r}
wb90 <- SMBassWB |>
  filter(yearcap==1990) |>
  select(-(species:gear),-(anu10:anu12))
peek(wb90,n=5)
```

&nbsp;

# Terminology
Two types of measurements can be made on calcified structures. A *radial* measurement is the total distance from the center of the structure (e.g., focus of scale or nucleus of otolith) to the edge of an annulus. An *incremental* measurement is the distance between two successive annuli. Radial measurements are required for back-calculation of fish length.

Back-calculation models estimate length at previous age $i$ (i.e., $L_{i}$) from known values of length at time of capture ($L_{C}$), scale radius to the $i$th annulus ($S_{i}$), and scale radius at time of capture ($S_{C}$). Several back-calculation models rely on the relationship between $S_{C}$ and $L_{C}$. Depending on the model, a function of mean $S_{C}$ for a given $L_{C}$ (i.e., $E(S_{C}|L_{C})$ ) or mean $L_{C}$ for a given $S_{C}$ (i.e., $E(L_{C}|S_{C})$) is used. These functions are not required to be linear, but often are, and in their linear form are represented as

$$ E(S_{C}|L_{C})=a+bL_{C} $$ {#eq-ESL}

$$ E(L_{C}|S_{C})=c+dS_{C} $$ {#eq-ELS}

Fitting these regressions is demonstrated [below](#scale-length-relationships).

&nbsp;

# Common Back-Calculation Models
The first back-calculation model was jointly developed by Knut Dahl and Einar Lea and appeared in @lea_methods_1910. The underlying concept of the *Dahl-Lea model* is that growth of the calcified structure is in exact proportion to growth in length of the fish. With this, the ratio of $S_{i}$ to $S_{C}$ is the same as the ratio of $L_{i}$ to $L_{C}$. Rearrangement of this equality yields the Dahl-Lea back-calculation model

$$ L_{i}=\frac{S_{i}}{S_{C}}L_{C} $$ {#eq-DahlLea}

The Dahl-Lea model describes a family of straight lines that pass through the **origin** and each observed ($S_{C},L_{C}$) point. Visually (@fig-BackCalcEx704), the estimated $L_{i}$ for a particular fish is found by locating $S_{i}$ along the x-axis, moving vertically until the straight line for that fish is met, and then moving horizontally to the point on the y-axis.

```{r}
#| label: fig-BackCalcEx704
#| fig-cap: Plot of length-at-capture versus scale radius for West Bearskin Lake Smallmouth Bass in 1990. All four methods of backcalculation are shown for fish 704 ($S_{2}$=3.49804, $L_{C}$=218, and $S_{C}$=7.44389; black point and line) with calculational steps shown with the arrows. Fish 701 is shown as the gray point and line for comparative purposes.
#| echo: FALSE
#| fig.width: 8
#| fig.height: 8

# ############################################################
# == BEGIN -- NOT SHOWN IN SUPPLEMENT, FOR PRINTING ONLY =====
par(mfrow=c(2,2),mar=c(3.05,3.05,1.15,0.65),mgp=c(1.5,0.3,0))
# ------------------------------------------------------------
# Helper functions
# ------------------------------------------------------------
makePlot <- function(name) {
  plot(lencap~radcap,data=wb90,xlim=c(0,max(radcap)),ylim=c(0,max(lencap)),
       xlab=expression(S[C]),ylab=expression(L[C]),yaxt="n",xaxt="n",cex=0.75)
  axis(2,at=seq(0,350,50),labels=c(0,50,NA,NA,seq(200,350,50)))
  xs <- c(0,2,6,8,10)
  axis(1,xs)
  mtext(name,line=0.1)
}
addCalc <- function(Si,Li,Sc,Lc,int) {
  points(Sc,Lc,pch=19)
  lines(c(0,Sc),c(int,Lc),lwd=2)
  text(Si,-10,paste(round(Si,3)),pos=1,xpd=TRUE)
  arrows(Si,-5,Si,Li,lwd=2,length=0.1,angle=20,xpd=TRUE)
  arrows(Si,Li,-0.1,Li,lwd=2,length=0.1,angle=20,xpd=TRUE)
  text(-0.1,Li,paste(round(Li,2)),pos=2,xpd=TRUE)
}

## Needed regressions and coefficients
lm.sl <- lm(radcap~lencap,data=wb90)
a <- coef(lm.sl)[1]; b <- coef(lm.sl)[2]
lm.ls <- lm(lencap~radcap,data=wb90)
c <- coef(lm.ls)[1]; d <- coef(lm.ls)[2]

## Info for fish #701
Sc701 <- 9.2219; Lc701 <- 312
## Info for fish #704
Sc <- 7.44389; Lc <- 218; Si <- 3.49804

# ------------------------------------------------------------
# Plots the Dahl-Lea
# ------------------------------------------------------------
makePlot("Dahl-Lea")
# Fish #701 as an example -- just show line
points(Sc701,Lc701,pch=19,col="grey",cex=1.25)
lines(c(0,Sc701),c(0,Lc701),lwd=2,col="grey")
# Fish #704 as an example -- show calculation
Li <- (Si/Sc)*Lc
addCalc(Si,Li,Sc,Lc,0)

# ------------------------------------------------------------
# Plots Fraser-Lee method
# ------------------------------------------------------------
makePlot("Fraser-Lee")
# Fish #701 as an example -- just show line
points(Sc701,Lc701,pch=19,col="grey",cex=1.25)
lines(c(0,Sc701),c(c,Lc701),lwd=2,col="grey")
# Fish #704 as an example -- show calculation
Li <- (Si/Sc)*(Lc-c)+c
addCalc(Si,Li,Sc,Lc,c)

#-------------------------------------------------------------------------------
# Plots the SPH method
#-------------------------------------------------------------------------------
makePlot("SPH")
# Fish #701 as an example -- just show line
points(Sc701,Lc701,pch=19,col="grey",cex=1.25)
lines(c(0,Sc701),c(-a/b,Lc701),lwd=2,col="grey")
# Fish #704 as an example -- show calculation
Li <- (-a/b)+(Lc+a/b)*(Si/Sc)
addCalc(Si,Li,Sc,Lc,-a/b)

#-------------------------------------------------------------------------------
# Plots the BPH method
#-------------------------------------------------------------------------------
makePlot("BPH")
# Fish #701 as an example -- just show line
int <- (c*Lc701)/(c+d*Sc701)
points(Sc701,Lc701,pch=19,col="grey",cex=1.25)
lines(c(0,Sc701),c(int,Lc701),lwd=2,col="grey")
# Fish #704 as an example -- show calculation
int <- (c*Lc)/(c+d*Sc)
Li <- Lc*(c+d*Si)/(c+d*Sc)
addCalc(Si,Li,Sc,Lc,int)
# == END -- NOT SHOWN IN SUPPLEMENT, FOR PRINTING ONLY =======
##############################################################
```

@fraser_growth_1916 was the first to describe, but @lee_review_1920 was the first to formally derive, the back-calculation model from the concept that "the growth *increment* of the scale is, on the average ..., a constant proportion of the growth *increment* of the fish" [@francis_back-calculation_1990]. In practice, the *Fraser-Lee model* modified the Dahl-Lea model by adjusting for the length of the fish when the calcified structure forms (i.e., $L=c$ when $S=0$), that is,

$$ L_{i}=\frac{S_{i}}{S_{C}}(L_{C}-c)+c $$ {#eq-FraserLee}

where $c$ comes from the length of the fish at the time of scale formation, the intercept of the length-scale relationship regression (e.g., from @eq-ELS), or from published "standards" for a species [@carlander_standard_1982]. The Fraser-Lee model describes a family of lines with an intercept of $c$ that pass through the ($S_{C},L_{C}$) point (@francis_back-calculation_1990; @fig-BackCalcEx704).

The *scale proportional hypothesis* (SPH) model was named by @francis_back-calculation_1990, but was first recognized by @whitney_interpretation_1956 when they said "*{i}f the scale was 10 per cent larger when the fish was caught than the average scale for that size of fish, [then] the scale would be 10 per cent larger than normal throughout the life.*"  If "average" and "normal" are considered to be expected values, then this hypothesis can be written as

$$ \frac{S_{i}}{E[S|L_{i}]}=\frac{S_{C}}{E[S|L_{C}]} $$

If it is assumed that the scale-length relationship is linear, then the two expected values in these ratios are computed by plugging $L_{i}$ and $L_{C}$, respectively, into the scale-length relationship (i.e., @eq-ESL) to produce

$$ \frac{S_{i}}{a+bL_{i}}=\frac{S_{C}}{a+bL_{C}} $$

which can be solved for $L_{i}$ to yield the general SPH back-calculation model

$$ L_{i} = \frac{S_{i}}{S_{C}}\left(L_{C}+\frac{a}{b}\right) - \frac{a}{b} $$ {#eq-SPH}

The linear SPH model produces a family of lines that all have an intercept of $-\frac{a}{b}$ and pass through each observed ($S_{C},L_{C}$) point (@fig-BackCalcEx704). The SPH model is the same as the Fraser-Lee model except that the intercept from @eq-ELS is replaced with $-\frac{a}{b}$. Further note that the SPH model is the same as the Dahl-Lea model if $a=0$.

The *body proportional hypothesis* (BPH) model was also named by @francis_back-calculation_1990 and was also first recognized by @whitney_interpretation_1956 when they said "*{i}f a fish at time of capture were 10 per cent smaller than the average fish with that size of scale, [then] the fish would be 10 per cent smaller than the expected length for the size of that scale throughout life.*" This hypothesis can be written as

$$ \frac{L_{i}}{E[L|S_{i}]}=\frac{L_{C}}{E[L|S_{C}]} $$

If the length-scale relationship is linear then the expected values can be found by plugging $S_{i}$ and $S_{c}$ into @eq-ELS to get

$$ \frac{L_{i}}{c+dS_{i}}=\frac{L_{C}}{c+dS_{C}} $$

which can be solved for $L_{i}$ to yield the general BPH back-calculation model

$$ L_{i}=L_{C}\frac{c+dS_{i}}{c+dS_{c}} $$ {#eq-BPH}

The linear BPH model produces a family of lines that have an intercept of $\frac{cL_{C}}{c+dS_{C}}$ and pass through each observed ($S_{C},L_{C}$) point (@fig-BackCalcEx704). In contrast to the other back-calculation models, the BPH model uses lines with a different intercept for each fish. The linear BPH model is the same as the Dahl-Lea model if $c=0$.

@vigliola_back-calculation_2009 described 18 other models for the back-calculation of fish length.^[Functions for each of these models can be created with `bcFuns()` from `RFishBC`.]

&nbsp;

# Data Organization & Manipulation
In *wide* or *one-fish-per-line* format, all information about a single fish, including all of the measurements from the calcified structure, is in one row of the data frame. `wb90` contains radial measurements in the wide format.^[See the data frame snippet [above](#data).]

For back-calculation of fish length, these data must be radial, and not incremental, measurements. If `wb90` had contained incremental measurements, then it could be converted to radial measurements with `gConvert()` from `RFishBC`. The `gConvert()` function requires the data frame with the incremental measurements as the first argument, the prefix (in quotes) for the columns that contain the incremental measurements in `in.pre=`, and the type of measurement to **convert to** in `out.type=` (the options are `"rad"` (the default) or `"inc"`). For example, the code below would create a new data frame from `wb90` with radial measurements (**IF** `wb90` had incremental measurements).

```{r}
#| eval: FALSE
# NOT RUN because wb90 is already radial measurements
wb90A <- RFishBC::gConvert(wb90,in.pre="anu",out.type="rad")
```

For efficient back-calculation, the data must also be converted to *long* or *one-measurement-per-line* format. Wide data may be converted to long data with `pivot_longer()` from `tidyr`.^[@ogleIntroductoryFisheriesAnalyses2016 used `gather()` from `tidyr` for this same purpose. The maintainers of `tidyr` suggest using `pivot_longer()` rather than `gather`.] For this use, the variables that contain the radial measurement data are given to `cols=`. The names for these variables will form a new column in the resulting data frame with the name given in `names_to=`. The values for these variables will be in a new column with the name given in `values_to=`.

```{r}
wb90r <- wb90 |>
  tidyr::pivot_longer(cols=anu1:anu9,names_to="agei",values_to="radi") |>
  arrange(fish,agei)
headtail(wb90r)
```

As noted in @ogleIntroductoryFisheriesAnalyses2016, there are three potential problems with this result. First, the new `agei` variable contains the names of the variables from the original wide format data frame (e.g., `anu1`, `anu2`) rather than numbers that correspond to the age that the annulus was formed. Converting these labels to numbers begins by finding the portion of the string after the "anu" prefix with `str_sub()` with the vector of names as the first argument and `start=4` (because "anu" is three characters long). The result from `str_sub()`, however, is a character that must then be converted to a numeric with `as.numeric()`.

The second problem is that several of the radial measurements contain `NA` values. The non-`NA` values are found and retained by using `!is.na()` within `filter()`.

The third problem, while not an issue with these particular data, is that "plus growth" may have been recorded. "Plus growth" is growth on the margin of the calcified structure that does not represent a complete year of growth. If "plus growth" is present, then the new `agei` variable will have a value greater than the age-at-capture value. These instances should be removed from the new long format data frame.

The following code adjusts for these three issues.

```{r}
wb90r <- wb90r |> 
  mutate(agei=as.numeric(stringr::str_sub(agei,start=4))) |>
  filter(!is.na(radi)) |>
  filter(agei<=agecap)
peek(wb90r,n=10)
```

&nbsp;

# Computing Back-Calculated Lengths
## Scale-Length Relationships
The scale-length (@eq-ESL) and length-scale (@eq-ELS) relationships required for all but the Dahl-Lea method are computed with the wide format data. Thus, the wide format data frame must contain the length of the fish (e.g., `lencap`) and the total radius of the calcified structure (e.g., `radcap`) at the time of capture. Both linear relationships are fit with `lm()` and the coefficients should be extracted with `coef()` and saved into objects.

```{r}
lm.sl <- lm(radcap~lencap,data=wb90)
( a <- coef(lm.sl)[[1]] )
( b <- coef(lm.sl)[[2]] )

lm.ls <- lm(lencap~radcap,data=wb90)
( c <- coef(lm.ls)[[1]] )
( d <- coef(lm.ls)[[2]] )
```

## Applying the Back-Calculation Models

The $L_{i}$ estimated with a back-calculation model are most easily added to the long format data frame with `mutate()`. For example, the $L_{i}$ computed with all four back-calculation models are added to `wb90r` below.

```{r}
wb90r <- wb90r |>
  mutate(DL.len=(radi/radcap)*lencap,
         FL.len=(radi/radcap)*(lencap-c)+c,
         SPH.len=(-a/b)+(lencap+a/b)*(radi/radcap),
         BPH.len=lencap*(c+d*radi)/(c+d*radcap))
headtail(wb90r,n=5)
```

For example, the mean length-at-age may be computed from the back-calculated lengths (shown below for the Fraser-Lee results).

```{r}
wb90r |>
  group_by(agei) |>
  summarize(n=validn(FL.len),mn=mean(FL.len),sd=sd(FL.len)) |>
  as.data.frame()
```

Additionally, the mean length at each back-calculated age computed separately for each age-at-capture may be computed with `sumTable()` from `FSA`, where the left side of the formula is the quantitative variable to be summarized and the right side has grouping variables presented in `row*column` format.

```{r}
sumTable(FL.len~agecap*agei,data=wb90r,digits=1)
```
