---
title: "Common Axis Labels on a Grid of Subplots"
author: "Derek H. Ogle"
output: html_document
---

Several methods for modifying base graphics were shown in @ogleIntroductoryFisheriesAnalyses2016, including simple layouts (side-by-side, one-over-the-other) of multiple plots. Creating common x- and y-axis labels for a grid of plots is described in this supplement. More complex layouts (e.g., plot in first row spans two columns or unequal sized plots) are [described in this supplement](complexLayouts.html).

A single function is used from `dplyr` in this supplement; thus, `::` is used instead of loading the entire package.

The `BullTroutRML2` data from `FSAdata` used in @ogleIntroductoryFisheriesAnalyses2016 are also used here.

```{r}
data(BullTroutRML2,package="FSAdata")
```

# Common Axis Labels on a Grid of Subplots
A common need is to plot multiple graphics in a grid-like format with one axis label that serves as the label for each graph. This can be accomplished with arguments to `par()`, but the regions of a plot must be understood first.

Each plot consists of three regions -- the plot area, the figure area, and the outer margin area (@fig-parEx1). The plot area is where the points or bars will be graphed. The area between the plot area and the figure area is where the axis ticks, labels, and title will appear. The area outside of the figure area is the outer margin area and is generally used to provide extra space around the graphic or to place text.

```{r}
#| label: fig-parEx1
#| echo: false
#| fig-cap: Schematic plot that illustrates the plotting area (inside the blue box), the figure area (inside the red box), and the outer margin area (between the dark gray and red boxes).
# ##############################################################################
# This code is not shown in the supplement, but is used to construct the plot
#   that demonstrates the three areas in a plot.
par(mar=c(1.5,1.5,0.5,0.5),mgp=c(3,1,0),oma=c(2,2,2,2))
X <- Y <- 0:10
plot(Y~X,type="n",xaxt="n",yaxt="n",xlab="",ylab="")

box("plot",col="blue")
text(5,5,"Plot Area",col="blue",cex=1.25)
box("figure",lwd=2,col="red")
mtext("Figure Area",side=1,line=0.2,adj=0.0,cex=1.25,col="red")
box("outer",lwd=3,col="gray50")
mtext("Outer Margin Area",side=3,line=0.4,adj=0.0,cex=1.25,col="gray50",outer=TRUE)
# ##############################################################################
```

In most instances (and the default), the width of the outer margin area is 0 on all sides of the figure area such that no outer margin area exists. The size of the outer margin area is set with `oma=` in `par()`, which takes a vector of four values to serve as widths of the four sides of the outer margin area, beginning with the bottom and moving counter-clockwise. The default is `par(oma=c(0,0,0,0))`. For example, margins that are two "lines" wide on the top and bottom and one "line" wide on the left and right may be set with `par(oma=c(2,1,2,1))`.

Common axis labels for multiple graphs can be placed in the outer margin area. In the example below, a 2-by-2 grid is created with `mfrow()` (as described in @ogleIntroductoryFisheriesAnalyses2016) and the bottom and left outer margin areas are increased to be two "lines" wide to allow for common x- and y-axis labels.

```{r}
#| label: SetPar
#| eval: false
par(mfrow=c(2,2),oma=c(2,2,0,0))
```

The four subplot areas are then populated with scatterplots (note that the x- and y-axis labels are set to empty strings to suppress labeling the axes)

```{r}
#| label: FillSubPlots
#| eval: false
xlmts <- c(-0.5,14.5)
ylmts <- c(0,700)
BTH1 <- BullTroutRML2 |> dplyr::filter(lake=="Harrison",era=="1977-80")
plot(fl~age,data=BTH1,main="Harrison, 1977-80",
     xlab="",ylab="",pch=19,xlim=xlmts,ylim=ylmts)
BTO1 <- BullTroutRML2 |> dplyr::filter(lake=="Osprey",era=="1977-80")
plot(fl~age,data=BTO1,main="Osprey, 1977-80",
     xlab="",ylab="",pch=19,xlim=xlmts,ylim=ylmts)
BTH2 <- BullTroutRML2 |> dplyr::filter(lake=="Harrison",era=="1997-01")
plot(fl~age,data=BTH2,main="Harrison, 1997-01",
     xlab="",ylab="",pch=19,xlim=xlmts,ylim=ylmts)
BTO2 <- BullTroutRML2 |> dplyr::filter(lake=="Osprey",era=="1997-01")
plot(fl~age,data=BTO2,main="Osprey, 1997-01",
     xlab="",ylab="",pch=19,xlim=xlmts,ylim=ylmts)
```

The common x- and y-axis labels are then placed in the outer margin areas with `mtext()`. In this capacity, `mtext()` requires the text to be written as the first argument, a number in `side=` that indicates the margin on which to print the text,^[The `side=` argument in `mtext()` uses the same numbering scheme as other functions -- 1=bottom, 2=left, 3=top, 4=right.] a number in `line=` that indicates the line on which to print the text (defaults to 0), and `outer=TRUE` to force the text into the outer margin area. Further note that `las=0` is used for the y-axis label to force the text to be parallel to the y-axis. This is only needed because `las=1` was set globally for producing this supplement.

```{r}
#| label: addmtext
#| eval: false
mtext("Age (years)",side=1,line=0,outer=TRUE,cex=1.3)
mtext("Fork Length (mm)",side=2,line=0,outer=TRUE,cex=1.3,las=0)
```

The combined code above creates the plot shown in @fig-layoutPar.

```{r}
#| label: fig-layoutPar
#| fig.cap: Grid of plots that uses the outer margin area to provide common axis labels.
#| fig.width: 8
#| fig.height: 8
#| echo: false
<<SetPar>>
par(mar=c(3.05,3.05,1,1),mgp=c(1.9,0.3,0),tcl=-0.2,
    las=1,cex.lab=0.95,cex.axis=0.9,cex.main=0.9)
<<FillSubPlots>>
<<addmtext>>
```
